Picky with peakpicking: assessing chromatographic peak quality with simple

### 2023-06-07 manuscript/manuscript.Rmd
Polishing up Fig 3 and surrounding text

We found that the predictions made from a Falkor-trained dataset consistently differed from a MESOSCOPE-trained dataset for the full and XCMS-only models. In the raw probability space, the two-parameter models had the highest $r$ value of 0.996, while the full models and the XCMS-trained models had $r$ values of 0.856 and 0.896, respectively. When compared in ranked space, we found an intensification of the effect described previously, with an even higher $r$ for the two-parameter model of 0.998 but lower $r$ values for the full and XCMS-trained model of 0.818 and 0.849, respectively (Figure 3).

@@

stepdown_cors <- dotplot_stepdown_data %>%
  group_by(training_set, which_params) %>%
  pivot_wider(names_from = training_set, values_from = c(pred_prob)) %>%
  summarize(pcor=cor(FT2040_features, MS3000_features, method = "pearson"),
            scor=cor(FT2040_features, MS3000_features, method = "spearman"))

raw_prob_space_gp <- dotplot_stepdown_data %>%
  pivot_wider(names_from = training_set, values_from = pred_prob) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(which_params=factor(
    which_params, levels=c("all_params", "xcms_params", "raw_data_params"),
    labels=c("All metrics", "XCMS metrics", "Raw data metrics"))) %>%
  ggplot() +
  geom_abline(slope = 1, intercept = 0, color="black", lwd=1) +
  geom_point(aes(x=FT2040_features, y=MS3000_features, color=feat_class), alpha=0.2) +
  facet_wrap(~which_params, nrow = 1) +
  scale_fill_manual(values = c("#a41118", "#f4bb23"), breaks = c("Bad", "Good"), 
                    aesthetics = c("color", "fill")) +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  coord_fixed() +
  labs(x="Falkor-trained model (predicted probability)", 
       y="MESO-trained model\n(predicted probability)") +
  theme_bw() +
  theme(legend.position = "none",
        strip.background = element_rect(fill="white"))
rank_prob_space_gp <- dotplot_stepdown_data %>%
  group_by(training_set, which_params) %>%
  mutate(pred_prob=rank(pred_prob)) %>%
  pivot_wider(names_from = training_set, values_from = pred_prob) %>%
  mutate(which_params=factor(
    which_params, 
    levels=c("all_params", "xcms_params", "raw_data_params"),
    labels=c("All metrics", "XCMS metrics", "Raw data metrics"))) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  ggplot() +
  geom_abline(slope = 1, intercept = 0, color="black", lwd=1) +
  geom_point(aes(x=FT2040_features, y=MS3000_features, color=feat_class), alpha=0.2) +
  facet_wrap(~which_params, nrow = 1) +
  scale_fill_manual(values = c("#a41118", "#f4bb23"), breaks = c("Bad", "Good"), 
                    aesthetics = c("color", "fill")) +
  scale_x_continuous(breaks = c(0, nrow(MS3000_features)), 
                     labels = c("Min rank", "Max rank")) +
  scale_y_continuous(breaks = c(0, nrow(MS3000_features)),
                     labels = c("Min\nrank", "Max\nrank")) +
  labs(x="Falkor-trained model (ranked likelihood)", 
       y="MESO-trained model\n(ranked likelihood)",
       color=NULL) +
  guides(color=guide_legend(override.aes = list(size=4, alpha=0.8))) +
  coord_fixed() +
  theme_bw() +
  theme(axis.text.x = ggtext::element_markdown(hjust=c(0, 1)),
        # strip.text = element_blank(),
        strip.background = element_rect(fill="white"),
        legend.position=c(0.875, 0.25),
        legend.background = element_blank(),
        legend.spacing.x = unit(0.003, units = "npc"))
comb_prob_space_gp <- egg::ggarrange(raw_prob_space_gp, rank_prob_space_gp, draw = FALSE)
ggsave(plot = comb_prob_space_gp, 
       filename = "comb_prob_space.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 5, units = "in", dpi = 300)
```

>>>>

We found that the predictions made from a Falkor-trained dataset consistently differed from a MESOSCOPE-trained dataset for the full and XCMS-only models. In the raw probability space, the two-parameter models had the highest $r$ value of 0.996, while the full models and the XCMS-trained models had $r$ values of 0.799 and 0.863, respectively. When compared in ranked space, we found an intensification of the effect described previously, with a higher $ρ$ for the two-parameter model of 0.998 but lower $ρ$ values for the full and XCMS-trained model of 0.725 and 0.804, respectively (Figure 3).

@@

stepdown_cors <- dotplot_stepdown_data %>%
  group_by(training_set, which_params) %>%
  pivot_wider(names_from = training_set, values_from = c(pred_prob)) %>%
  summarize(pcor=cor(FT2040_features, MS3000_features, method = "pearson"),
            scor=cor(FT2040_features, MS3000_features, method = "spearman")) %>% 
  mutate(pcor=paste0("r=", signif(pcor, digits = 3))) %>%
  mutate(scor=paste0("ρ=", signif(scor, digits = 3))) %>%
  mutate(which_params=factor(
    which_params, levels=c("all_params", "xcms_params", "raw_data_params"),
    labels=c("All metrics", "XCMS-only metrics", "Raw data metrics")))

raw_prob_space_gp <- dotplot_stepdown_data %>%
  pivot_wider(names_from = training_set, values_from = pred_prob) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(which_params=factor(
    which_params, levels=c("all_params", "xcms_params", "raw_data_params"),
    labels=c("All metrics", "XCMS-only metrics", "Raw data metrics"))) %>%
  ggplot() +
  geom_abline(slope = 1, intercept = 0, color="black", lwd=1) +
  geom_point(aes(x=FT2040_features, y=MS3000_features, color=feat_class), alpha=0.2) +
  geom_text(aes(label=pcor), x=-Inf, y=Inf, hjust=-0.1, vjust=1.5, data=stepdown_cors) +
  facet_wrap(~which_params, nrow = 1) +
  scale_fill_manual(values = c("#a41118", "#f4bb23"), breaks = c("Bad", "Good"), 
                    aesthetics = c("color", "fill")) +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  coord_fixed() +
  labs(x="Falkor-trained model (predicted probability)", 
       y="MESO-trained model\n(predicted probability)") +
  theme_bw() +
  theme(legend.position = "none",
        strip.background = element_rect(fill="white"))
rank_prob_space_gp <- dotplot_stepdown_data %>%
  group_by(training_set, which_params) %>%
  mutate(pred_prob=rank(pred_prob)) %>%
  pivot_wider(names_from = training_set, values_from = pred_prob) %>%
  mutate(which_params=factor(
    which_params, 
    levels=c("all_params", "xcms_params", "raw_data_params"),
    labels=c("All metrics", "XCMS-only metrics", "Raw data metrics"))) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  ggplot() +
  geom_abline(slope = 1, intercept = 0, color="black", lwd=1) +
  geom_point(aes(x=FT2040_features, y=MS3000_features, color=feat_class), alpha=0.2) +
  geom_text(aes(label=scor), x=-Inf, y=Inf, hjust=-0.1, vjust=1.5, data=stepdown_cors) +
  facet_wrap(~which_params, nrow = 1) +
  scale_fill_manual(values = c("#a41118", "#f4bb23"), breaks = c("Bad", "Good"), 
                    aesthetics = c("color", "fill")) +
  scale_x_continuous(breaks = c(0, nrow(MS3000_features)), 
                     labels = c("Min rank", "Max rank")) +
  scale_y_continuous(breaks = c(0, nrow(MS3000_features)),
                     labels = c("Min\nrank", "Max\nrank")) +
  labs(x="Falkor-trained model (ranked likelihood)", 
       y="MESO-trained model\n(ranked likelihood)",
       color=NULL) +
  guides(color=guide_legend(override.aes = list(size=4, alpha=0.8))) +
  coord_fixed() +
  theme_bw() +
  theme(axis.text.x = ggtext::element_markdown(hjust=c(0, 1)),
        # strip.text = element_blank(),
        strip.background = element_rect(fill="white"),
        legend.position=c(0.925, 0.25),
        legend.background = element_blank(),
        legend.spacing.x = unit(0.003, units = "npc"))
comb_prob_space_gp <- egg::ggarrange(raw_prob_space_gp, rank_prob_space_gp, draw = FALSE)
ggsave(plot = comb_prob_space_gp, 
       filename = "comb_prob_space.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 5, units = "in", dpi = 300)
```
